---
title: "R and RStudio Self-Test"
author: "Prepared by: Victoria Field"
date: "Summer 2020"
output:
  html_document:
    toc: TRUE
    toc_float: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Purpose** This self-test is designed to assess your current skills in R and RStudio. 

**Acknowledgements** Major components of this document are directly derived from lessons hosted at [DataCarpentry](https://datacarpentry.org/), Wickham and Grolemund's [R for Data Science](https://r4ds.had.co.nz/index.html), and 

# Before we begin

Be sure to download both [R](https://www.r-project.org/) and [RStudio](https://rstudio.com/)

**What's the difference?**

`R` is a language. `RStudio` is an *integrated development environment* (IDE) that allows you to code in R, edit, annotate, plot, and organize. To learn more about the differences and why they're important, click [here](https://datacarpentry.org/R-ecology-lesson/00-before-we-start.html#what_is_r_what_is_rstudio)

## Assumptions

For the purposes of this self-test, we will assume you know how to generate and execute commands in the `R` language. Ideally: 

- you are working in `RStudio`

- you understand the difference between the 4 panes (if not, see [this](https://datacarpentry.org/R-ecology-lesson/00-before-we-start.html#knowing_your_way_around_rstudio))
  
  - **Console**
  
  - **Source**
  
  - **Environment/History/Other**
  
  - and **File/Plots/Packages/Help/Viewer**
  
- you can easily find and describe your working directory (if not, see [this](https://r4ds.had.co.nz/workflow-projects.html))

- you understand file pathways (if not, see [this](https://r4ds.had.co.nz/workflow-projects.html))

## Most importantly, remember

Learning a new skill, especially if you're learning from scratch, can be fun and empowering! However, it can also be challenging, frustrating, and down-right aggravating. If this is your first time working with a coding language, it can be overwhelming (and that's putting it mildly!). The good news is that no matter where you're starting from, there is an abundance of resources to help you along the way, and you can take comfort in the fact that others have been where *you* are, and have gotten to where *they* are by not giving up! Take breaks as needed, ask questions early and often, create a workflow that works for you, and constantly be on the look-out for ways to improve your scripts, workflow, and version control (more on those later!)

# Begin Self-Test

Here you will be presented with a series of "challenges". Some scipts (codes), packages, and functions will be provided to you and you will be asked to use them to answer questions. After each challenge, a link to more information will be provided so that you can learn more about the skill tested and check your answers. 

## Introduction to R

**Challenge #1:** What are the final values of `mass`, `age`, and `mass_index`
```{r}
mass <- 47.5            
age  <- 122             
mass <- mass * 2.0      
age  <- age - 20        
mass_index <- mass/age  
```

For more information about creating, re-writing, manipulating, and viewing objects, click [here](https://datacarpentry.org/R-ecology-lesson/01-intro-to-r.html#creating_objects_in_r) 

**Challenge #2a:** What type of objects are `num_char`, `num_logical`, `char_logical`, and `tricky`?
```{r}
num_char <- c(1, 2, 3, "a")
num_logical <- c(1, 2, 3, TRUE)
char_logical <- c("a", "b", "c", TRUE)
tricky <- c(1, 2, 3, "4")
```

**Challenge #2b:** How many values in `combined_logical` are "TRUE" (character). Why? If there was a coercion, where did it happen?
```{r}
num_logical <- c(1, 2, 3, TRUE)
char_logical <- c("a", "b", "c", TRUE)
combined_logical <- c(num_logical, char_logical)
```

For more information about vectors and data types, click [here](https://datacarpentry.org/R-ecology-lesson/01-intro-to-r.html#vectors_and_data_types)

**Challenge 3a:** Using the object `heights` created below, create a new object called `heights_no_na` with no `NA` values.
```{r}
heights <- c(63, 69, 60, 65, NA, 68, 61, 70, 61, 59, 64, 69, 63, 63, NA, 72, 65, 64, 70, 63, 65)
```
For more information about how to handle missing data, click [here](https://datacarpentry.org/R-ecology-lesson/01-intro-to-r.html#missing_data)

**Challenge 3b:** What is the median height of the new `heights_no_na` vector?

**Challenge 3c:** How many of the heights from `heights_no_na` are greater than 67?

**Challenge 3d:** From the `heights` vector, what is the 3rd and 8th value?

For more information on subsetting vectors, click [here](https://datacarpentry.org/R-ecology-lesson/01-intro-to-r.html#subsetting_vectors)

## Starting with data

**Challenge 1a** Download a dataset using the `download.file()` function. The code below is an example. You must supply the correct file path for the `destfile=` argument
```{r}
download.file(url="https://ndownloader.figshare.com/files/2292169",
              destfile = "./data_raw/portal_data_joined.csv")
```

**Challenge 1b:** Bring the downloaded data file into the R environment using the `read.csv` function. The code below is an example. You must supply the correct file path
```{r}
surveys<-read.csv("./data_raw/portal_data_joined.csv")
```

**Challenge 2:** 

1. What type of object is `surveys`?

2. How many rows? How many columns?

3. How many species have been recorded? 

For more information about dataframes, click [here](https://datacarpentry.org/R-ecology-lesson/02-starting-with-data.html#what_are_data_frames)

**Challenge 3:** Create an object called `surveys_middle` using the `nrow()` function that contains only the *middle* row of data from `surveys`

For more information on indexing and subsetting a dataframe in base `R`, click [here](https://datacarpentry.org/R-ecology-lesson/02-starting-with-data.html#indexing_and_subsetting_data_frames)

**Challenge 4:** 

1. Create a new object called `sex` from `surveys$sex`

2. What are the levels of the `sex`? 

3. Rename the `" "` level of `sex` to `"undetermined"`

4. Rename `"M"` and `"F"` to `"male"` and `"female"`, respectively. 

5. Create a basic barplot of `sex` where the level `"undetermined"` is plotted last (i.e., at the right-most position)

For more information on factors, levels, re-naming, and re-ordering, click [here](https://datacarpentry.org/R-ecology-lesson/02-starting-with-data.html#factors)

**Challenge 5:** 

1.Use the package `lubridate` to create add a Date column to `surveys`

2. There should be 129 NAs in your new Date column. Identify the rows of `year`, `month`, and `day` that are failing to parse into the Date column. Why are they failing?

For more information on working with dates in `R`, click [here](https://datacarpentry.org/R-ecology-lesson/02-starting-with-data.html#formatting_dates) or [here](https://r4ds.had.co.nz/dates-and-times.html)

## Manipulating Data 

**Challenge 1:** Using pipes, function `filter`, and function `select` from the `dplyr` package, subset the surveys data to include animals collected before 1995 and retain only the columns `year`, `sex`, and `weight`.

For more information on pipes, click [here](https://datacarpentry.org/R-ecology-lesson/03-dplyr.html#pipes) or [here](https://www.datacamp.com/community/tutorials/pipe-r-tutorial)

**Challenge 2:** Use pipes, `group_by()`, and `summarize()` to find the mean, min, and max hindfoot length for each species (using `species_id`). Also add the number of observations (hint: see `?n`).

For more information the *split-apply-combine* paradigm, click [here](https://datacarpentry.org/R-ecology-lesson/03-dplyr.html#split-apply-combine_data_analysis_and_the_summarize()_function)

**Challenge 3:** 

1. The surveys data set has two measurement columns: `hindfoot_length` and `weight`. This makes it difficult to do things like look at the relationship between mean values of each measurement per year in different plot types. First, use `gather()` to create a dataset where we have a key column called `measurement` and a value column that takes on the value of either `hindfoot_length` or `weight`. Hint: You’ll need to specify which columns are being gathered.

2. With this new data set, calculate the average of each measurement in each `year` for each different `plot_type`. Then `spread()` them into a data set with a column for `hindfoot_length` and `weight`. Hint: You only need to specify the key and value columns for `spread()`

**Challenge 4:**

1. Create a new dataframe called `surveys_complete` that has the following characteristics: 

- no missing values for `weight`, `hindfoot_length`

- no observations for `sex: undetermined`

2. Create a dataframe called `species_counts` that extracts the most common `species_id`s from `surveys_complete` (common = more than **50** observations)

3. Manipulate `surveys_complete` to include only those species found in `species_counts`

For more information on reshaping data, click [here](https://datacarpentry.org/R-ecology-lesson/03-dplyr.html#reshaping_with_gather_and_spread)

**Challenge 5:** Write your new `surveys_complete` dataframe to file and be sure to supply the correct path